---
title: "Extended example"
format: html
---

Typically, package functions include multiple arguments, the different values ​​of which allow for flexible configuration. Below are various examples demonstrating the package's customization potential.

## Set up

Import the package.

```{r}
library(paneldesc)
```

## Data import

Import the built-in dataset with simulated unbalanced panel data.

```{r}
data(production)
```

Set up a panel structure in advance so that you don't have to do it later each time you use other functions.
Note that if `delta` is supplied, the function checks for omitted time periods. If such periods exist, they will be taken into account when other functions work with this argument.

```{r}
panel <- make_panel(production, index = c("firm", "year"), delta = 1)
```

## Panel data structure analysis

If necessary, you can display more detailed statistics on the distribution of entities by periods and periods by entities.

```{r}
describe_balance(panel, detail = TRUE, digits = 2)
```

`plot_periods()` function allows to customize colors. 

```{r}
plot_periods(panel, colors = c("gray", "black"))
```

Both `describe_patterns()` and `plot_patterns()` allow you to limit the range of patterns to display. In addition, `desribe_patterns()` allows to customize rounding when calculting shares, while `plot_patterns()` allows to customize colors. 

```{r}
describe_patterns(panel, limits = 3, digits = 2)
```

```{r}
plot_patterns(panel, limits = c(4, 6), colors = c("darkgray", "white"))
```

## Missing values analysis

`plot_missing()` allows to select specific variables to analyze. One can also customize colors. 

```{r}
plot_missing(
  panel,
  select = c("labor", "capital"),
  colors = c("black", "white")
)
```

`summarize_missing()` also allows to select specific variables to analyze. In addition, it can provide more detailed table by adding period-specific missing value counts. One can also customize rounding. 

```{r}
summarize_missing(
  panel,
  select = c("labor", "capital"),
  detail = TRUE,
  digits = 2
)
```

`describe_incomplete()` can provide more detailed table by adding variable-specific missing value counts.

```{r}
describe_incomplete(panel, detail = TRUE)
```

## Numeric variables analysis

`summarize_numeric()` allows to select specific variables to analyze. It also can provide additional statisitcs. Custom rounding is also available. 

```{r}
summarize_numeric(
  panel,
  select = c("sales", "labor"),
  group = "year",
  detail = TRUE,
  digits = 2
)
```

`plot_heterogeneity()` allows to choose several grouping variables. One can also customize colors. 

```{r}
plot_heterogeneity(
  panel,
  select = "sales",
  group = c("firm", "industry", "year"),
  colors = c("black", "gray")
)
```

`decompose_numeric()` allows to select specific varibles to analyze. One also can output the resulting table in wide format and hide redundant details, as well as customize rounding. 

```{r}
decompose_numeric(
  panel,
  select = c("sales", "labor"),
  detail = FALSE,
  format = "wide",
  digits = 2
)
```

`plot_demeaned()` allows to choose several grouping variables. One can also customize colors. 

```{r}
plot_demeaned(
  production,
  select = "sales",
  group = c("industry", "region", "year"),
  colors = c("gray", "black")
)
```

```{r}
plot_demeaned(
  production,
  select = c("labor", "sales"),
  group = c("industry", "region", "year"),
  colors = c("gray", "black")
)
```


## Factor variables analysis

Compared to the default settings, both `decompose_factor()` and `summarize_transition()` functions can output the resulting table in long format and also allow to customize rounding. `descompose_factor()` also allows to select specific variables to analyze.

```{r}
decompose_factor(panel, select = "industry", format = "long", digits = 2)
```

```{r}
summarize_transition(panel, select = "industry", format = "long", digits = 2)
```