---
title: "For Stata users"
format: html
---

In this example, the package functions are combined in such a way that the results obtained are more or less familiar to those users who are accustomed to using `Stata`.

## Set up

Import the package.

```{r}
library(paneldesc)
```

## Data import

Import the built-in dataset with simulated unbalanced panel data.

```{r}
data(production)
```

## summarize

You can look at the simple descriptive statistics with `summarize_numeric()`.

```{r}
summarize_numeric(production)
```

## xtset

You can set up a panel structure in advance so that you don't have to do it later each time you use other functions. Note that if `delta` is supplied, the function checks for omitted time periods. If such periods exist, they will be taken into account when other functions work with this argument.

```{r}
panel <- make_panel(production, index = c("firm", "year"), delta = 1)
```

## xtdes

`xtdes` command funcitonality can be reproduced by combining `describe_dimensions()`, `describe_balance()`, and `describe_patterns()` functions.

```{r}
describe_dimensions(panel)
```

```{r}
describe_balance(panel, detail = TRUE)
```

```{r}
describe_patterns(panel)
```

## xtsum

`xtsum` command funcitonality can be reproduced with `decompose_numeric()` function.

```{r}
decompose_numeric(panel)
```

## xttab

`xttab` command funcitonality can be reproduced with `decompose_factor()` function.

```{r}
decompose_factor(panel)
```

## xttrans

`xttrans` command funcitonality can be reproduced with `summarize_transition()` function.

```{r}
summarize_transition(panel, select = "industry")
```

## reshape

`reshape` command funcitonality can be reproduced with `make_wide()` and `make_long()` functions. Unlike the Stata command, these functions do not require you to specify all variables to transform (all variables are transformed by default), but they do allow to specify time-invariant variables, if any.

```{r}
wide <- make_wide(panel, static = "region")
names(wide)
```

```{r}
long <- make_long(wide, static = "region")
names(long)
```

## xtdata

`xtdata` command funcitonality can be partially reproduced with `make_demeaned()` function. Particularly, `xtdata, fe` is reproduced by passing only firm identifier to the function. 

```{r}
firm_demeaned <- make_demeaned(production, group = "firm")
```

Unlike Stata command, `make_demeaned()` allows you to demean variables not only within entities, but using various grouping patterns. For example, when data frame has panel attributes, it will be demeaned using both entities and time identifiers. Grouping by more than two variables is also available. 

```{r}
firm_year_demeaned <- make_demeaned(panel)
```

`xtdata, be` and `xtdata, re` functionality is not reproduced directly. However, with some additional efforts, `xtdata, be` can be replicated with `make_mundlak()` and `summarize_numeric()` as both these functions allow to calculate within-group means. 