---
title: "For Stata users"
format: html
---

In this example, the package functions are combined in such a way that the results obtained are more or less familiar to those users who are accustomed to using `Stata`.

## Set up

Import the package.

```{r}
library(paneldesc)
```

## Data import

Import the built-in dataset with simulated unbalanced panel data.

```{r}
data(production)
```

## summarize

You can look at the simple descriptive statistics with `summarize_numeric()`.

```{r}
summarize_numeric(production)
```

## xtset

You can set up a panel structure in advance so that you don't have to do it later each time you use other functions. Note that if `delta` is supplied, the function checks for omitted time periods. If such periods exist, they will be taken into account when other functions work with this argument.

```{r}
panel <- make_panel(production, index = c("firm", "year"), delta = 1)
```

## xtdes

`xtdes` command funcitonality can be reproduced by combining `describe_dimensions()`, `describe_balance()`, and `describe_patterns()` functions.

```{r}
describe_dimensions(panel)
```

```{r}
describe_balance(panel, detail = TRUE)
```

```{r}
describe_patterns(panel)
```

## xtsum

`xtsum` command funcitonality can be reproduced with `decompose_numeric()` function.

```{r}
decompose_numeric(panel)
```

## xttab

`xttab` command funcitonality can be reproduced with `decompose_factor()` function.

```{r}
decompose_factor(panel)
```

## xttrans

`xttrans` command funcitonality can be reproduced with `summarize_transition()` function.

```{r}
summarize_transition(panel, select = "industry")
```