---
title: "Comparison with other packages"
format: html
---

Almost all of this package's functionality can be replicated using individual functions from other packages. However, I believe the main advantage of the `paneldesc` package is that it combines all the functions necessary for panel data analysis and implements them in a single format. Below is a comparison with individual functions from other packages.

## Set up

Import packages to compare.

```{r, message=FALSE}
library(paneldesc)
library(tidyr)
library(gplots)
library(modelsummary)
library(plm)
library(fixest)
library(pdynmc)
library(xtsum)
```

## Data import

Import the built-in dataset with simulated unbalanced panel data.

```{r}
data(production)
```

Create unbalanced panel for the analysis purposes. 

```{r}
unbalanced <- production[rowSums(!is.na(production[, c("sales", "capital", "labor", "industry")])) > 0, ]
```

## paneldesc vs. pdynmc

`paneldesc::plot_patterns()` and `pdynmc::strucUPD.plot()` visualize the participation patterns of entities in an unbalanced panel. `pdynmc::strucUPD.plot()` groups similar participation patterns by the number of periods and separates different patterns by bar color, while preserving the original order of the entities. In turn, `paneldesc::plot_patterns()` reorders entities by their patterns to show the dominance of each pattern using bar width.

```{r}
paneldesc::plot_patterns(unbalanced, index = c("firm", "year"))
```

```{r}
pdynmc::strucUPD.plot(unbalanced, i.name = "firm", t.name = "year")
```

As can be seen above, the two functions produce similar results for an apparently unbalanced panel. However, applying `pdynmc::strucUPD.plot()` to the original `production` data frame results in an error, since it is technically balanced (the same number of years for each firm), while `paneldesc::plot_patterns()` detects entity-time pairs without other data and produces the same plot, showing the unbalanced panel structure.

```{r}
paneldesc::plot_patterns(production, index = c("firm", "year"))
```

## paneldesc vs. modelsummary

Among many other functions, `paneldesc::summarize_numeric()` and `modelsummary::datasummary()` provide tables with descriptive statistics for numeric variables and also allow the user to calculate statistics within groups. The main advantage of `modelsummary::datasummary()` is the flexibility of the output appearance, allowing various parameters to be specified for the table structure and output format. In turn, `paneldesc::summarize_numeric()` always returns a data frame, preserving the user's freedom of further work with it. Furthermore, in my opinion, the syntax of `paneldesc::summarize_numeric()` is simpler compared to `modelsummary::datasummary()`, although this is a matter of taste.

```{r}
paneldesc::summarize_numeric(
  production, 
  select = c("capital", "labor"), 
  group = "year"
)
```

```{r}
modelsummary::datasummary(
  Factor(year) * (capital + labor) ~ N + Mean + SD + Min + Max, 
  data = production
)
```

## paneldesc vs. gplots

The functions `paneldesc::plot_heterogeneity()` and `gplots::plotmeans()` allow the user to analyze heterogeneity between and within groups. The results of these functions are quite similar: they visualize group means and the within-group variance of the selected variable. The main difference is that `gplots::plotmeans()` displays group size labels and plots confidence intervals, while `paneldesc::plot_heterogeneity()` plots a point for each observation to show both the variance and the group size.

```{r}
paneldesc::plot_heterogeneity(production, select = "sales", group = "year")
```

```{r}
gplots::plotmeans(sales ~ year, production)
```

## paneldesc vs. xtsum

`paneldesc::decompose_numeric()` and `xtsum::xtsum()` provide descriptive statistics for numeric variables in panel data. Specifically, they decompose the standard deviation into between-value and within-value components, mimicking the `xtsum` command in Stata. The syntax and behavior of the functions are quite similar. By default, `xtsum::xtsum()` returns output in HTML format, although it allows customization. In turn, `paneldesc::summarize_numeric()` always returns a data frame, preserving the user's freedom for further manipulation.

```{r}
paneldesc::decompose_numeric(
  production, 
  select = c("sales", "capital", "labor"), 
  index = "firm"
)
```

```{r}
xtsum::xtsum(
  production, 
  variables = c("sales", "capital", "labor"), 
  id = "firm", 
  t = "year", 
  na.rm = TRUE
)
```

## paneldesc vs. plm

`paneldesc::make_balanced()` and `plm::make.pbalanced()` allow the user to obtain various types of balanced panel datasets. 

```{r}
paneldesc_bal_1 <- paneldesc::make_balanced(
  unbalanced, 
  index = c("firm", "year"), 
  balance = "entities"
)
dim(paneldesc_bal_1)
```

```{r}
paneldesc_bal_2 <- paneldesc::make_balanced(
  unbalanced, 
  index = c("firm", "year"), 
  balance = "periods"
)
dim(paneldesc_bal_2)
```

```{r}
paneldesc_bal_3 <- paneldesc::make_balanced(
  unbalanced, 
  index = c("firm", "year"), 
  balance = "rows"
)
dim(paneldesc_bal_3)
```

```{r}
plm_bal_1 <- plm::make.pbalanced(
  unbalanced, 
  index = c("firm", "year"), 
  balance.type = "shared.individuals"
)
dim(plm_bal_1)
```

```{r}
plm_bal_2 <- plm::make.pbalanced(
  unbalanced, 
  index = c("firm", "year"), 
  balance.type = "shared.times"
)
dim(plm_bal_2)
```

```{r}
plm_bal_3 <- plm::make.pbalanced(
  unbalanced, 
  index = c("firm", "year"), 
  balance.type = "fill"
)
dim(plm_bal_3)
```

As can be seen above, the two functions produce similar results for an apparently unbalanced panel. However, applying `plm::make.pbalanced()` to the original `production` dataframe results in no changes, as it is technically balanced (the same number of years for each firm), while `paneldesc::make_balanced()` automatically detects entity-time pairs without other data, revealing the unbalanced panel structure. This allows the function to balance the panel in the same way as shown above.

```{r}
paneldesc_bal_1b <- paneldesc::make_balanced(
  production, 
  index = c("firm", "year"), 
  balance = "entities"
)
dim(paneldesc_bal_1b)
```

```{r}
paneldesc_bal_2b <- paneldesc::make_balanced(
  production, 
  index = c("firm", "year"), 
  balance = "periods"
)
dim(paneldesc_bal_2b)
```

```{r}
paneldesc_bal_3b <- paneldesc::make_balanced(
  production, 
  index = c("firm", "year"), 
  balance = "rows"
)
dim(paneldesc_bal_3b)
```

```{r}
plm_bal_1b <- plm::make.pbalanced(
  production, 
  index = c("firm", "year"), 
  balance.type = "shared.individuals"
)
dim(plm_bal_1b)
```

```{r}
plm_bal_2b <- plm::make.pbalanced(
  production, 
  index = c("firm", "year"), 
  balance.type = "shared.times"
)
dim(plm_bal_2b)
```

```{r}
plm_bal_3b <- plm::make.pbalanced(
  production, 
  index = c("firm", "year"), 
  balance.type = "fill"
)
dim(plm_bal_3b)
```

## paneldesc vs. fixest

`paneldesc::make_demeaned()` and `fixest::demean()` allow the user to demean variables  using various grouping patterns. As can be seen below, basic applications of these functions to a single variable produce similar results. When working with multiple variables, the behavior of the functions differs slightly. `fixest::demean()` is aimed at regression analysis and therefore removes all rows with missing values ​​for at least one variable and performs calculations based on the resulting dataset. `paneldesc::make_demeaned()`, in contrast, preserves missing values ​​in the dataset and performs transformations on the full dataset (when calculating within-group means, missing values ​​are removed for each variable separately, which often allows for more data to be used compared to `fixest::demean()`). Thus, the `paneldesc::make_demeaned()` function may be preferable for exploratory data analysis, as it preserves more observations. In turn, `fixest::demean()` is more flexible in configuration, allowing the user to select many additional parameters.

```{r}
demeaned_paneldesc <- paneldesc::make_demeaned(
    production[, c("sales", "firm", "year")], 
    group = c("firm", "year")
)
summary(demeaned_paneldesc)
```

```{r}
demeaned_fixest <- fixest::demean(
  X = production[, "sales"],
  f = production[, c("firm", "year")]
)
summary(demeaned_fixest)
```

## paneldesc vs. tidyr

Among other functions, `paneldesc::make_wide()` and `paneldesc::make_long()`, as well as `tidyr::pivot_wider()` and `tidyr::pivot_longer()`, allow the user to transform panel data from long to wide format and from wide to long, respectively. The key advantage of the `paneldesc` functions is their simpler and more automatic implementation in a panel data context, including the ability to handle time-varying and static variables separately. In turn, the `tidyr` functions are more adaptable for a broader range of applications, although they lack automatic solutions for panel data. For example, the `region` variable in the `production` dataset is a time-invariant variable, but has missing observations due to the unbalanced structure of the panel data. In this case, the `tidyr` functions require data preprocessing, which somewhat complicates their implementation (omitted here for brevity). Also, I find the syntax of the `paneldesc` function simpler than `tidyr`, although that's a matter of taste.

```{r}
wide_paneldesc <- paneldesc::make_wide(
    production, 
    select = c("sales", "capital", "labor", "industry", "ownership"),
    index = c("firm", "year")
)
```

```{r}
long_paneldesc <- paneldesc::make_long(
    wide_paneldesc, 
    select = c("sales", "capital", "labor", "industry", "ownership"), 
    index = c("firm", "year")
)
```

```{r}
wide_tidyr <- tidyr::pivot_wider(
  production,
  id_cols = firm,
  names_from = year,
  values_from = !c(firm, year)
)
names(wide_tidyr)
```

```{r}
long_tidyr <- tidyr::pivot_longer(
  wide_tidyr,
  cols = !firm,
  names_to = c(".value", "year"),
  names_pattern = "(.*)_(.*)"
)
names(long_tidyr)
```